### 2012-10-07 selected-results.tex
Adding some selected (available) results (selected-results.pdf combines some of the available graphs)
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@812 df1a90cd-247b-4364-bdaf-287a704eb499

\documentclass[onecolumn]{IEEEtran}
% \include{graphicx}
\usepackage{pdfpages}
\usepackage{hyperref}

\begin{document}

\tableofcontents

\clearpage

\textbf{\color{red}Left graphs show bandwidths of data/interset stream\\
Right graphs show the same data but in number of packets}

\section{Large Tree}

\begin{figure}[h]
  \centering

\includegraphics[scale=1]{selected-results/topo-large-tree.pdf}

  \caption{128-node binary tree topology}
\end{figure}

The following presents results for three independent runs of the simulation, varying number of good and bad clients.

\clearpage

\subsection{\textbf{4 evil nodes}, 60 good nodes}

\subsubsection{Dynamic Limits, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/dynamic-topo-large-tree-evil-4.pdf}
\end{minipage}

\clearpage

\subsubsection{Probabilistic Interest accept, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/probabilistic-topo-large-tree-evil-4.pdf}
\end{minipage}

% \clearpage

% \subsubsection{Physical Limits, Run 1-3}

% \begin{minipage}[b]{\textwidth}
% \includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/simple-limits-topo-large-tree-evil-4.pdf}
% \end{minipage}

% \clearpage

% \subsubsection{No Limits, Run 1-3}

% \begin{minipage}[b]{\textwidth}
% \includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/no-limits-topo-large-tree-evil-4.pdf}
% \end{minipage}

\clearpage
\subsection{\textbf{8 evil nodes}, 56 good nodes}

\subsubsection{Dynamic Limits, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.85,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/dynamic-topo-large-tree-evil-8.pdf}
\end{minipage}

\clearpage

\subsubsection{Probabilistic Interest accept, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/probabilistic-topo-large-tree-evil-8.pdf}
\end{minipage}

\clearpage
\subsection{\textbf{12 evil nodes}, 52 good nodes}

\subsubsection{Dynamic Limits, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.85,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/dynamic-topo-large-tree-evil-12.pdf}
\end{minipage}

\clearpage

\subsubsection{Probabilistic Interest accept, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/probabilistic-topo-large-tree-evil-12.pdf}
\end{minipage}

\clearpage
\subsection{\textbf{16 evil nodes}, 48 good nodes}

\subsubsection{Dynamic Limits, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.85,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/dynamic-topo-large-tree-evil-16.pdf}
\end{minipage}

\clearpage

\subsubsection{Probabilistic Interest accept, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/probabilistic-topo-large-tree-evil-16.pdf}
\end{minipage}

\clearpage
\subsection{\textbf{20 evil nodes}, 44 good nodes}

\subsubsection{Dynamic Limits, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.85,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/dynamic-topo-large-tree-evil-20.pdf}
\end{minipage}

\clearpage

\subsubsection{Probabilistic Interest accept, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/probabilistic-topo-large-tree-evil-20.pdf}
\end{minipage}

\clearpage
\subsection{\textbf{24 evil nodes}, 40 good nodes}

\subsubsection{Dynamic Limits, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.85,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/dynamic-topo-large-tree-evil-24.pdf}
\end{minipage}

\clearpage

\subsubsection{Probabilistic Interest accept, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/probabilistic-topo-large-tree-evil-24.pdf}
\end{minipage}

\clearpage
\subsection{\textbf{28 evil nodes}, 36 good nodes}

\subsubsection{Dynamic Limits, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.85,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/dynamic-topo-large-tree-evil-28.pdf}
\end{minipage}

\clearpage

\subsubsection{Probabilistic Interest accept, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/probabilistic-topo-large-tree-evil-28.pdf}
\end{minipage}

% \clearpage
% \subsection{\textbf{32 evil nodes}, 32 good nodes}

% \subsubsection{Dynamic Limits, Run 1-3}

% \begin{minipage}[b]{\textwidth}
% \includepdf[scale=0.85,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/dynamic-topo-large-tree-evil-32.pdf}
% \end{minipage}

% \clearpage

% \subsubsection{Probabilistic Interest accept, Run 1-3}

% \begin{minipage}[b]{\textwidth}
% \includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/probabilistic-topo-large-tree-evil-32.pdf}
% \end{minipage}

\clearpage

\textbf{\color{red}Left graphs show bandwidths of data/interset stream\\
Right graphs show the same data but in number of packets}

\section{Tree}

\begin{figure}[h]
  \centering

\includegraphics[scale=1]{selected-results/topo-tree.pdf}

  \caption{32-node binary tree topology}
\end{figure}

The following presents results for three independent runs of the simulation, varying number of good and bad clients.

\clearpage
\subsection{\textbf{1 evil node}, 15 good nodes}

\subsubsection{Dynamic Limits, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/dynamic-topo-tree-evil-1.pdf}
\end{minipage}

\clearpage

\subsubsection{Probabilistic Interest accept, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/probabilistic-topo-tree-evil-1.pdf}
\end{minipage}

\clearpage

\subsubsection{Physical Limits, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/simple-limits-topo-tree-evil-1.pdf}
\end{minipage}

% \clearpage

% \subsubsection{No Limits, Run 1-3}

% \begin{minipage}[b]{\textwidth}
% \includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/no-limits-topo-tree-evil-1.pdf}
% \end{minipage}

\clearpage
\subsection{\textbf{2 evil nodes}, 14 good nodes}

\subsubsection{Dynamic Limits, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.85,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/dynamic-topo-tree-evil-2.pdf}
\end{minipage}

\clearpage

\subsubsection{Probabilistic Interest accept, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/probabilistic-topo-tree-evil-2.pdf}
\end{minipage}

\clearpage

\subsubsection{Physical Limits, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/simple-limits-topo-tree-evil-2.pdf}
\end{minipage}

\clearpage
\subsection{\textbf{3 evil nodes}, 13 good nodes}

\subsubsection{Dynamic Limits, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.85,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/dynamic-topo-tree-evil-3.pdf}
\end{minipage}

\clearpage

\subsubsection{Probabilistic Interest accept, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/probabilistic-topo-tree-evil-3.pdf}
\end{minipage}

\clearpage

\subsubsection{Physical Limits, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/simple-limits-topo-tree-evil-3.pdf}
\end{minipage}

\clearpage
\subsection{\textbf{4 evil nodes}, 12 good nodes}

\subsubsection{Dynamic Limits, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.85,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/dynamic-topo-tree-evil-4.pdf}
\end{minipage}

\clearpage

\subsubsection{Probabilistic Interest accept, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/probabilistic-topo-tree-evil-4.pdf}
\end{minipage}

\clearpage

\subsubsection{Physical Limits, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/simple-limits-topo-tree-evil-4.pdf}
\end{minipage}

\clearpage
\subsection{\textbf{5 evil nodes}, 11 good nodes}

\subsubsection{Dynamic Limits, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.85,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/dynamic-topo-tree-evil-5.pdf}
\end{minipage}

\clearpage

\subsubsection{Probabilistic Interest accept, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/probabilistic-topo-tree-evil-5.pdf}
\end{minipage}

\clearpage

\subsubsection{Physical Limits, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/simple-limits-topo-tree-evil-5.pdf}
\end{minipage}

\clearpage
\subsection{\textbf{6 evil nodes}, 10 good nodes}

\subsubsection{Dynamic Limits, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.85,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/dynamic-topo-tree-evil-6.pdf}
\end{minipage}

\clearpage

\subsubsection{Probabilistic Interest accept, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/probabilistic-topo-tree-evil-6.pdf}
\end{minipage}

\clearpage

\subsubsection{Physical Limits, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/simple-limits-topo-tree-evil-6.pdf}
\end{minipage}

\clearpage
\subsection{\textbf{7 evil nodes}, 9 good nodes}

\subsubsection{Dynamic Limits, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.85,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/dynamic-topo-tree-evil-7.pdf}
\end{minipage}

\clearpage

\subsubsection{Probabilistic Interest accept, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/probabilistic-topo-tree-evil-7.pdf}
\end{minipage}

\clearpage

\subsubsection{Physical Limits, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/simple-limits-topo-tree-evil-7.pdf}
\end{minipage}

% \clearpage
\subsection{\textbf{8 evil nodes}, 8 good nodes}

\subsubsection{Dynamic Limits, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.85,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/dynamic-topo-tree-evil-8.pdf}
\end{minipage}

\clearpage

\subsubsection{Probabilistic Interest accept, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/probabilistic-topo-tree-evil-8.pdf}
\end{minipage}

\clearpage

\subsubsection{Physical Limits, Run 1-3}

\begin{minipage}[b]{\textwidth}
\includepdf[scale=0.9,nup=2x3,pages=1-6]{selected-results/near-tree-root-bandwidhts-and-packets/simple-limits-topo-tree-evil-8.pdf}
\end{minipage}

\end{document}