### 2012-10-23 algos/dynamic-limits.tex
Adding fair-queueing algorithm description
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@815 df1a90cd-247b-4364-bdaf-287a704eb499

% \subsection{Dynamic limits}

\subsection{Dynamic limits assignment algorithm}
\label{sec:dynamic limits}

% At the same time, 
% semi-guarantee

% that there will not be congestion, while fully utilizing downstream link.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../paper"
%%% End: