### 2012-11-27 large-scale.tex
added simulation vs emulation
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@823 df1a90cd-247b-4364-bdaf-287a704eb499

\subsection{Large scale simulations}
\label{sec:largescale}

Large scale, large scale

%\begin{figure}[htpb]
%  \centering
%  \includegraphics[scale=0.5]{figures/sim-emu-power.pdf}
%  \caption{Strength of Interest flooding attack}
%  \label{fig:simemupower}
%\end{figure}

%\begin{figure}[htpb]
%  \centering
%  \includegraphics[scale=0.5]{figures/sim-emu-performance.pdf}
%  \caption{Data retrieval by legitimate clients}
%  \label{fig:simemuperf}
%\end{figure}