### 2012-11-29 ddos-ndn.tex
restruturing
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@831 df1a90cd-247b-4364-bdaf-287a704eb499

\section { DDoS Attacks in NDN \label{ddos-ndn}}
Brief overview of possible attacks in NDN, and scope of what we will tackle in this paper (Interest flooding).

%\begin{figure}[htpb]
%  \centering
%  \includegraphics[scale=0.5]{figures/sim-emu-power.pdf}
%  \caption{Strength of Interest flooding attack}
%  \label{fig:simemupower}
%\end{figure}

%\begin{figure}[htpb]
%  \centering
%  \includegraphics[scale=0.5]{figures/sim-emu-performance.pdf}
%  \caption{Data retrieval by legitimate clients}
%  \label{fig:simemuperf}
%\end{figure}