### 2012-12-02 design.tex
Adding a couple of figures, more to come
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@843 df1a90cd-247b-4364-bdaf-287a704eb499

\section{Design}
\label{sec:design}

\subsection{Interest satisfaction statistics}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=4.5cm]{trivial-topo}
  \caption{Trivial topology (all links are 3Mbps/10ms)}
  \label{fig:trivial topology}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=1]{limits}
  \caption{Dynamics of the unsatisfied Interests statistics on gateway's interface towards the attacker}
  \label{fig:ratio example}
\end{figure}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: