### 2012-12-02 large-scale.tex
Adding figures (topology sketch and simulation results) for large scale
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@845 df1a90cd-247b-4364-bdaf-287a704eb499

\subsection{Large scale simulations}
\label{sec:largescale}

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.26,angle=90]{7018-r0}
  \caption{Large-scale topology: 344 client routers (red), 109 gateway routers (green), 109 backbone routers (blue)}
  \label{fig:large-scale-topo}
\end{figure}

\begin{figure*}[t]
 \centering
 \includegraphics[scale=1]{topo-7018-gw-15mins-new/7018-r0-good-0-producer-gw}
 \caption{Satisfaction ratio dynamics during the attack (30\% attackers)}
 % producer on a gateway node
 \label{fig:large-scale}
\end{figure*}

% \begin{figure*}[t]
%  \centering
%  \includegraphics[scale=1]{topo-7018-bb-15mins-new/7018-r0-good-0-producer-bb}
%  \caption{Satisfaction ratio dynamics during the attack (30\% attackers)}
%  % producer on a backbone node
%  \label{fig:large-scale}
% \end{figure*}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: