### 2012-12-02 04-design.tex

\section{Design of Interest flooding mitigation methods}
\label{sec:design}

% \section {Countermeasures against Interest flooding}

% Basic limits, probablity based on statistics, and dynamic window adjustment

% (Description of each of the above, and their implications)

% \input{design}

\subsection{Interest satisfaction statistics}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=4.5cm]{trivial-topo}
  \caption{Trivial topology (all links are 3Mbps/10ms)}
  \label{fig:trivial topology}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=1]{limits}
  \caption{Dynamics of the unsatisfied Interests statistics on gateway's interface towards the attacker}
  \label{fig:ratio example}
\end{figure}

\input{algos/physical-limits.tex}

\input{algos/fair-queueing.tex}

\input{algos/dynamic-limits.tex}

\input{algos/probabilistic.tex}

\input{algos/stats.tex}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: