### 2012-12-02 07-discussion.tex

\section{Discussion}
\label{sec:discussion}

Lessons we learnt, comparison of different mitigation techniques. 
Adapting these to mitigate other DDoS attacks.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: