### 2012-12-03 04-design.tex
Correcting algorithms. Far far far from perfect...
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@849 df1a90cd-247b-4364-bdaf-287a704eb499

\section{Interest flooding mitigation methods}
\label{sec:design}

% Points what should be here:
% - What can be done (in general) to mitigate flooding attack
% - Which building blocks NDN architecture gives to mitigate Interest flooding attacks: Interest limits, ability to measure Interest satisfaction performance (Interest satisfaction stats)
% - Methods to set these limits: static/dynamic
% - Methods how limits can be applied: best-effort (fifo), "fair" queuing, probabilistic
% - Reference to caching: we don't consider it here, but caching provides additional level of protection, especially for certain types of attacks

% Our definition of attack mitigation is that good clients are still able to access data on the producer.

In this section we present several methods to mitigate Interest flooding attacks in NDN, featuring different degrees of implementation complexity, as well as different degrees of effectiveness.

The presented methods rely on the core NDN principle of the flow balance between Interest and Data packets: Data packet can be forwarded only in response to the incoming Interest and each Interest is satisfied by at most one Data packet.
As a result of this principle, all traffic on NDN is ``receiver''-driven, where the ``receiver'' is either a user or any router on the path between the user and a Data producer.
Therefore, any node can limit the number of forwarded Interests, effectively limiting the amount of the retrieved Data.
At the same time, each forwarded Interest can be used to build up various data plane performance statistics, such as per-incoming interface ratios of satisfied Interests.

% Combining per-incoming interface data plane performance statistics with

% Disclaimer
% The presented solutions to mitigate Interest flooding DDoS attacks

% In particular, we explore different mechanisms to app

% Basic limits, probablity based on statistics, and dynamic window adjustment

\input{algos/physical-limits.tex}

\input{algos/stats.tex}

\input{algos/fair-queueing.tex}

\input{algos/probabilistic.tex}

\input{algos/dynamic-limits.tex}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: