### 2012-12-12 04b-design-intelligent.tex

\subsection{Intelligent attack mitigation}
\label{sec:intelligent mitigating}

While the Interest limit is the key building block to suppress mechanisms of the Interest flooding attack, legitimate Interests need to be somehow prioritized and malicious Interests need to be somehow penalized in order to completely suppress the attack.
That is, instead of processing Interests always based on the first-in-first-serve rule, NDN routers need some basis to treat the incoming Interests differently.
Thus, the primary task in bringing intelligence to the Interest flooding attack mitigation is to proactively distinguish between legitimate and malicious Interests.

To achieve such an Interest differentiation we can leverage another unique feature of the stateful forwarding in NDN---namely, the guaranteed symmetric flow of Interest and Data packets.
In other words, NDN guarantees that a forwarded Interest would always results in receiving a corresponding Data, provided such Data exists and Interest/Data packets were not lost on the way.
Therefore, all Interests that result in Data return can be considered legitimate, while the ones that always timeout should be deemed as malicious.\footnote{Recall that in order to maximize effect of the Interest flooding attack, an adversary expresses a large volume of junk Interests (see Section~\ref{sec:interest flooding}).  Implications of other types of attacks are discussed in Section~\ref{sec:discussion}.}

Unfortunately, the timeout-based differentiation is reactive by nature: one cannot know in advance that an Interest would timeout or bring Data back.
What we can do proactively is to keep an up-to-date statistics of Interests satisfaction ratios (number of forwarded versus number of satisfied Interests), and use this statistics to project outcomes for new Interests.
For example, maintaining independent Interest satisfaction ratio statistics for each incoming interface can be enough to estimate whether an Interest received from a neighbor will bring Data or time out.
Statistics can also be kept at more granular levels (e.g., per outgoing interface, per prefix, etc.), which can further improve the estimation quality.

The devil is always in the details.
From the one hand, such statistics needs to start penalizing adversaries as soon as possible (i.e., negative stats should build up fast).
On the other hand, the positive statistics should not deteriorate too fast (i.e., positive stats should be relatively long-term).
Our preliminary experiments showed that the standard exponentially weighted moving average, performed once a second with $\alpha$ coefficient $e^{-1/30}$, approximately corresponding to a 30-second averaging window, provides a good balance between the two contradictory requirements.

% \subsubsection{\textbf{Data plane performance tracking}}
% \label{sec:stats}

\input{algos/stats.tex}

%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{\textbf{Probabilistic Interest accept}}
\label{sec:probabilistic}

\input{algos/probabilistic.tex}

%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{\textbf{Dynamic limits}}
\label{sec:dynamic limits}

\input{algos/dynamic-limits.tex}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: