### 2012-12-17 paper.tex
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@884 df1a90cd-247b-4364-bdaf-287a704eb499

% \documentclass[10pt, conference, compsocconf]{IEEEtran}
\documentclass[conference]{IEEEtran}
% http://networking2013.poly.edu/call-for-papers/paper-submission/
% Page limit 9 pages

\makeatletter
\def\ps@headings{%
\def\@oddhead{\mbox{}\scriptsize\rightmark \hfil \thepage}%
\def\@evenhead{\scriptsize\thepage \hfil \leftmark\mbox{}}%
\def\@oddfoot{}%
\def\@evenfoot{}}
\makeatother

\pagestyle{headings}

\title{Mitigating Interest Flooding DDoS Attacks in NDN}%:\\ Cache/Content Poisoning}
\author{anonymous}

\newfont{\nicettfont}{cmtt10}
\newcommand{\ndnName}[1]{``{\nicettfont #1}''}
\renewcommand{\texttt}[1]{{\nicettfont #1}}
\newcommand{\todo}[1]{\vspace{2 mm}\par \noindent \marginpar{\textsc{ToDo}}
\framebox{\begin{minipage}[c]{0.95 \columnwidth}
\tt #1 \end{minipage}}\vspace{5 mm}\par}

\usepackage{graphicx}
% \usepackage[colorlinks]{hyperref}
\usepackage[]{hyperref}
\usepackage{breakurl}
\usepackage{url}
\usepackage[nocompress]{cite}
% \usepackage{verbatim}
% \usepackage{algpseudocode}
\usepackage{algpseudocode,algorithm}
% More customizeable version. Probably it would be better to convert pseudocode to 2e format
% \usepackage{algorithm2e} 
\usepackage{multirow}

\graphicspath{{figures/}}

\begin{document}

\maketitle

\begin{abstract}
Distributed Denial of Service attack is an ongoing problem in today's Internet, where packets from large numbers of compromised hosts thwart the paths to the victim site and/or overload the victim machines. 
In a newly proposed future Internet architecture, Named Data Networking (NDN), end users request desired data by sending Interest packets, and the network forwards data packets upon request only, effectively eliminating the existing DDoS attacks. 
However an NDN network can be subject to a new type of DoS attacks, namely Interest packet flooding.  
In this paper we investigate effective solutions to mitigate Interest flooding.
We show that NDN's inherent properties of each router keeping per packet state and flow-balance (i.e., one Interest packet retrieves at most one Data packet) provides the  basis for effective DDoS mitigation algorithms.
Our evaluation through simulations shows that the solution can quickly and effectively respond to Interest flooding.
\end{abstract}

\begin{IEEEkeywords}
Information-Centric Networks, Named-data Networking, Denial-of-Service
\end{IEEEkeywords}

\input{01-intro}

\input{02-ndn-background}

\input{03-interest-flooding}

\input{04a-design-naive}

\input{04b-design-intelligent}

\input{05-evaluation}

\input{07-discussion}

\input{06-related}

\input{08-conclusion}

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}