### 2012-12-18 04a-design-naive.tex
changes to 4a physical limits
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@886 df1a90cd-247b-4364-bdaf-287a704eb499

\section{Interest flooding mitigation methods}
\label{sec:design}

% Points what should be here:
% - What can be done (in general) to mitigate flooding attack
% - Which building blocks NDN architecture gives to mitigate Interest flooding attacks: Interest limits, ability to measure Interest satisfaction performance (Interest satisfaction stats)
% - Methods to set these limits: static/dynamic
% - Methods how limits can be applied: best-effort (fifo), "fair" queuing, probabilistic
% - Reference to caching: we don't consider it here, but caching provides additional level of protection, especially for certain types of attacks

% Our definition of attack mitigation is that good clients are still able to access data on the producer.

In this section we present several algorithms to mitigate Interest flooding attacks in NDN.  Our mitigation strategies feature varying degrees of implementation complexity and effectiveness -- the higher the implementation complexity, the more effective is the algorithm against Interest flooding attacks. We start by describing two naive strategies and use the insights and lessons learnt from the evaluation of these techniques to inform and design more effective mitigation techniques that work well in a variety of attack scenarios.

%   - Naive approach to Interest Limits (that is called physical limits everywhere else)
%     Example how this can be implemented in a simple way
%     Baseline solution
%   - Interest limits with "fair" queuing
%     Improving problem of simple limits (no single face can dominate), but doesn't solve the proble

% - Attack mitigation
%   - Per-incoming interface Interest statistics

%   - Dynamic Interests limit adjustments

%   - Probabilistic Interest accept

Our mitigation strategies are based on two fundamental principles of NDN architecture - flow balance between Interest and Data packets, and Data packets traversing the reverse path taken by Interest packets. The flow balance in NDN refers to the fact that one Interest packet can be satisfied with at most one Data packet. This principle allows intermediate routers that forward the Interest packets to control the number of outstanding Interests packets in the network. The fact that each Data packet takes the reverse path of the corresponding Interest packet allows intermediate routers to match a received Data packet to the corresponding Interest packet and thus determine which Interest packets were satisfied. As we describe in the rest of this section,  NDN routers can exploit this information to make informed decisions on which Interest packets to forward, how many Interest packets to forward, and thus effectively defend against DDoS attacks.

%Our methods to mitigate Interest flooding attack rely on the fundamental principle of the NDN architecture: the flow balance between Interest and Data packets: one Interest packet (the only communication initiator) can be satisfied with at most one Data packet.
%Because NDN is host-to-host architecture (as opposed to end-to-end in the current IP Internet), the flow balance principle allow any entity on the network, end-hosts and routers, control what and how much Data they want to receive.
%Therefore, any node can limit the number of forwarded Interests, effectively limiting the amount of the retrieved Data.
%At the same time, each forwarded Interest can be used to build up various data plane performance statistics, such as per-incoming interface ratios of satisfied Interests.

%\subsection{Na\"{i}ve attack mitigation}

%We start with a couple of naive strategies that can be used to migitate attacks in an NDN network and describe the pitfalls associated with these strategies.

% In a normal network operation, the size of Interests packets is supposed to be significantly smaller than the size of the requested Data.

% All communication in NDN network is host-to-host and receiver-driven.

An obvious and simple strategy that can be used to mitigate the Interest flooding attack is for a router to limit the number of forwarded Interests out of each interface based on the physical capacity of the corresponding out interface. This technique is a slight modification of the well-known {\it Token Bucket} algorithm that is currently widely used in packet switched networks. Analogous to the {\it Token Bucket} algorithm, NDN routers keep track of the amount of data requested that can fully utilize the downstream link (calculated from the number of pending Interests) and once the link capacity limit has been reached, they no longer forward new incoming Interests. Ideally, the number of tokens {\it i.e.} the \emph{Pending Interests Limit} for each link will be proportional to the link's bandwidth-delay product (BDP)~\cite{tcp-survey}. We can formalize this value as follows:
% With the objective to request as many Data packets, as downstream link can pump through, we are getting the following equation for Interest limit:

\begin{equation}
\mathrm{Interest\ Limit} = Delay\ [s] \cdot \frac{\mathrm{Bandwidth\ [Bytes/s]}}{\mathrm{Data\ packet\ size\ [Bytes]}}
\end{equation}

In the above equation, \emph{Delay} is the expected time of Interest being satisfied and \emph{Data packet size} is the size of the returned Data packet.
Although both these values are not known {\it apriori}, it is not really necessary to use their exact values.
One can simply set the pending Interest limit based on the average values of round trip time and observed Data packet size, as the network buffers can smooth out most of the network fluctuations.

An obvious failure of the above {\it Token Bucket} algorithm is that it can be used as an effective mechanism to carry out denial of service attacks. If a router has utilized all its {\it tokens} to forward {\it bad} Interests on an out Interface, it can no longer forward incoming Interests from legitimate users on that out inteface till the pending {\it bad} Interests start to expire. One way to get around this issue is to impose a per-interface limit so that {\it bad} Interests are not allowed to entirely consume the limits of a specific interface. We describe this technique in greater detail in the next subsection.

%That is, if it is known that the amount of already requested data ($=$~amount of forwarded Interests) can fully utilize the downstream link, an NDN node---either an end-user or an NDN router---has absolutely no point in forwarding new incoming Interests and creating corresponding PIT entries.
%For example, if a router A on Fig.~\ref{fig:flooding example} already forwarded 125 Interests requesting 1000-byte Data packets, any new incoming Interests can be almost safely dropped, provided the link capacity between A--B is 100~Mbps and delay is 10~ms.
%In other words, 125 Data packets returned by router B will fully utilize the link ($125 \times 1000 \mathrm{~bytes} \approx 10\mathrm{~ms} \times 100\mathrm{~Mbps}$) and any excess Data would be dropped.
%
%\subsubsection{\textbf{Interest limits (physical limits)}}
%\label{sec:physical limits}
%
%\input{algos/physical-limits.tex}

%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{\textbf{Physical limits with per-interface fairness}}
\label{sec:queuing}

\input{algos/fair-queueing.tex}

%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%