### 2013-01-13 06-related.tex
A new version of related work. May need corrections
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@902 df1a90cd-247b-4364-bdaf-287a704eb499

\section{Related work}
\label{sec:related work}

The Named Data Networking is a relatively new communication paradigm, but it already got a limited attention in the security community.
Lauinger~\cite{Lauinger:2010:Security--scalability} showed several attack vectors that can be applied to NDN. However, the work is specifically limited to a particular design choices implemented in CCNx software (\url{http://www.ccnx.org/}), rather than NDN architecture itself.  
W\"ahlisch et al.~\cite{Wahlisch:2012:Backscatter-from} explore security and stability threats in the general area of information-centric networks (ICN), using CCNx software as an example.  
While this work attempts to generalize findings for all ICNs, it largely does not go beyond the current design choices of CCNx.
% (which is an example, but not the final word in NDN) 
% Gasti et al.~\cite{gasti2012ddos} identified several NDN architecture-specific attacks (Interest flooding and content/cache poisoning) and proposed potential solutions, without devising particular algorithms.
In this paper we follow some suggestions by Gasti et al.~\cite{gasti2012ddos} and explore intrinsic potential of NDN architecture to sustain DDoS attacks, using the Interest flooding attack as a first step in this direction.
Our design for the attack mitigation relies in part on limiting number of outstanding Interests.  
Similar techniques were explored by Yi et al.~\cite{adaptive-forwarding,adaptive-tr} and Rozhnova and Fdida~\cite{rozhnova2012effective}, but applied only to facilitate performance in the NDN network.

The general area of denial of service attacks and their mitigation methods is one of the hottest topics of the past decade.
% Among all the body of work, 
The most related area to this paper is solutions to network-level flooding attacks, which include DDoS detection techniques~\cite{Chen:2007:Collaborative-detection, Jun:2011:DDoS-flooding}, methods to pushback of malicious traffic to the edged using various filtering techniques~\cite{Pushback, Tupakula:2003:A-practical-method, Argyraki:2005:Active-internet, Oikonomou:2006:A-framework-for-a-collaborative, Liu:2008:To-filter-or-to-authorize:, Chou:2009:Proactive-surge, Liu:2010:NetFence:-preventing}, various ticketing systems with conditional admission of traffic to the network~\cite{Yaar:2004:SIFF:-A-stateless, Yang:2005:A-DoS-limiting-network, Wendlandt:2006:Fastpass:-Providing, Natu:2007:Fine-grained-capabilities, Portcullis, Capabilities}, overlay-based filtering~\cite{Stone:2000:CenterTrack:-An-IP-overlay, Keromytis:2004:SOS:-An-architecture, Kline:2011:Shield:-DoS-filtering}, systems that attempt to use approximate IP traffic symmetry to estimate and filter out malicious traffic~\cite{Wang:2002:Detecting-SYN-flooding, Kreibich:2005:Using-packet, Mahimkar:2007:dFence:-Transparent}.

% Our design for the Interest flooding attack mitigation borrows

% This paper talks about attacks on content centric infrastructure (not specifically to NDN).
% \cite{Wahlisch:2012:Backscatter-from}

% % Collaborative detection of DDoS attacks over multiple network domains
% \cite{Chen:2007:Collaborative-detection} distributed change-point detection

% % DDoS flooding attack detection through a step-by-step investigation
% \cite{Jun:2011:DDoS-flooding}

% % A practical method to counteract denial of service attacks
% \cite{Tupakula:2003:A-practical-method}

% Alex: We should not include this reference for the review, but afterwards
% Previous work on DDoS and DoS~\cite{gasti2012ddos}

% Network level (Flooding)
% % 2002
% Pushback~\cite{Pushback} (hop-by-hop push back) aggregate- based congestion control/congestion signatures (e.g., per-destination), rate-limiting, explicit notification
% AITF~\cite{Argyraki:2005:Active-internet} (active filtering requests; path recording, last-point of trust detecting)?
% DefCOM \cite{Oikonomou:2006:A-framework-for-a-collaborative} Communication between src, core, and dst to provide collaborative defense. Dst raises alerts, src differentiate bad/good, packet tagging. Core rate-limit, congestion-aware re-tag. Traffic prioritization.
% StopIt \cite{Liu:2008:To-filter-or-to-authorize:} (filtering) tagging with passports, tagged prioritization, dst- invoked filter instantiation
% NetFence \cite{Liu:2010:NetFence:-preventing} congestion policing: fair queuing (one leaky bucket per {src,L}, L- bottleneck), secure "ticketing" with signals for AIMD to increase/decrease bandwidth allowance
% PSP \cite{Chou:2009:Proactive-surge} topology as origin-destination pairs, packet tagging, OD traffic isolation, sharing based on historic bandwidth demand

% SIFF~\cite{Yaar:2004:SIFF:-A-stateless} (time-based tickets, capabilities)
% TVA \cite{Yang:2005:A-DoS-limiting-network} like SIFF, traffic- based tickets, hierarchical per- AS, per-source fair queuing
% FastPass \cite{Wendlandt:2006:Fastpass:-Providing} (tokens: various puzzles, other methods) traffic authorization, routers verify and filter.
% \cite{Natu:2007:Fine-grained-capabilities} based on TVA/SIFF: specifies previously unspecified mechanisms to grant "tickets" for requesters, extends capabilities to include priorities, changing tickets to be only destination- dependent, not path dependent
% % LazySusan \cite{Crowcroft:2007:Lazy-Susan:} TR (latency-based proof of work--- computational puzzles)
% Portcullis \cite{Portcullis} Protecting connection setup with computational proof of work, tickets and traffic prioritization for the rest, per- computational fairness; DNS to disseminate puzzles
% There are many other puzzle- based schemes, on which Portcullis is based or which problems it solves

% PacketSymmetry \cite{Kreibich:2005:Using-packet} (leverages packet symmetry to detect and remove bad traffic) packet asymmetry - number of transmitted packets (tx) to received packets (rx); negative values when rx outweighs tx, positive values when tx outweighs rx, and zero in the case of perfectly balanced traffic)
% % Detecting SYN flooding attacks
% \cite{Wang:2002:Detecting-SYN-flooding} based on the protocol behavior of TCP SYN-FIN (RST) pairs, and is an instance of the Seqnential Change Point Detection
% dFence \cite{Mahimkar:2007:dFence:-Transparent} (middleboxes) TCP connection proxying, rate estimation and TCP/ACK modification

% Overlays:
% CenterTrack \cite{Stone:2000:CenterTrack:-An-IP-overlay} (overlay) reroute suspicious traffic to a special routers with diagnostic capabilities using an overlay
% SOS \cite{Keromytis:2004:SOS:-An-architecture}
% Shield \cite{Kline:2011:Shield:-DoS-filtering}

% NDN and other future information centric architectures were lacking any DDoS mitigation techniques, so we put efforts in understanding DDoS defensive systems, which were made for IP architecture, for the purpose of possible incorporation of them in NDN. In this section we will overview generic directions of DDoS mitigation designs and try to think about which tricks could be used in NDN.

% Capability-based systems allow routers to negotiate, perform and enforce limitations on bandwidth consumption on router-to-router and router-to-client links. Usually such systems require an extensive trust infrastructure in order to validate secure keys used by routers and clients. This implies that each client must pass through an authentication process prior to using any bandwidth~\cite{Capabilities}. We believe that for NDN this method doesn't fit really well for the following reasons. First, this method is very questionable in terms of scalability and mobility support since it requires tight trust connections between routers and hosts. Thus, it neglects one of the main advantages (mobility) of NDN. Second, the design of trust infrastructure for NDN is not well defined yet, which means that we cannot really make assumptions in our design of mitigation techniques.

% Computation-based systems provide each client an access to the network resources only after performing a significant computations such as solving puzzles. Spending a huge amounts of computational resources can effectively slow down a flooding attack by the botnet, however it also creates additional computational burden for legitimate clients. One of the main difficulties in such systems is the question of puzzle dessimination. In ~\cite{Portcullis} authors suggested to propagate puzzles using DNS system. In NDN since routers have caching abilities, we can propagate puzzles just like a normal data packets, using Interest-Data exchange. Overall, this method looks interesting and deserves being investigated in a future work.

% Congestion control systems do per flow traffic analysis and drop of packets belonging to misbehaving flows. For instance, Random Early Detect (RED)~\cite{RED} identifies flows that do not comply with TCP-friendly end-to-end congestion control, and preferentially drop them. Such techniques, if largely deployed, could perform well, however, they cannot provide effective defense against non-greedy botnets that are creating a huge amount of low-bandwidth flows. Since NDN routers maintain a richer per-packet state, which translates into a better per-flow statistics, these techniques can become much more effective when applied to NDN.

% Push back systems are trying to detect bad and good traffic flows on each router, and once attack is detected by one of the routers, it starts a coordinated push back by downrating incoming flows with a bad traffic in order to provide more capacity for a good traffic. This process is reiterating downstream till it reaches edge routers, which are directly connected to attacking bot machines~\cite{Pushback}. In general this technique can be applied to NDN architecture thanks to its decentralized coordination of downrating bad traffic flows. Moreover, NDN provides better capabilities for accurate flow detection.

%were looking for getting accustomed

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: